Precision calculations of the cosmic shear power spectrum projection

### 2018-01-17 limber2.tex
eq 1 round brackets
Removed marked comments according to ref report

\title[Precision calculations of the cosmic shear projection]{Precision calculations of the cosmic shear power spectrum projection}

@@

We compute the spherical-sky weak-lensing power spectrum of the shear and
convergence. We discuss various approximations, such as flat-sky, and first- and
second-order Limber equations for the projection. We find that the impact of
adopting these approximations \forref{is} negligible when constraining cosmological
parameters from current weak lensing surveys.
%
This is demonstrated using data from the Canada-France-Hawaii Telescope Lensing Survey
(CFHTLenS). We find that the reported tension with Planck Cosmic Microwave
Background (CMB) temperature anisotropy results cannot be alleviated.
%, in
%contrast to the recent claim made by Kitching et al. (2016, version 1)

%\citet{2016arXiv161104954K}.

For future large-scale surveys with unprecedented precision, we show that the
spherical second-order Limber approximation will provide sufficient accuracy.
In this case, the cosmic-shear power spectrum is shown to be in agreement with
the full projection at the sub-percent level for $\ell > 3$, with the
corresponding errors an order of magnitude below cosmic variance for all
$\ell$. When computing the two-point shear correlation function, we show that
the flat-sky fast Hankel transformation results in errors below two percent
compared to the full spherical transformation.
%
In the spirit of reproducible research, our numerical implementation of all
approximations and the full projection are publicly available within the
package \textsc{nicaea} at \texttt{http://www.cosmostat.org/software/nicaea}.

@@

In the era of the upcoming large-scale surveys that will provide measurements
of cosmic shear with unprecedented precision, one needs to revisit the
theoretical predictions of the observables to ensure that the accuracy of the
models meet the accuracy of the observations. In this paper, we examine the
widely used Limber approximation for the projected weak-lensing power spectrum.
We consider spherical coordinates and the flat-sky approximation, and compute
the full projection \forref{of the lensing power spectrum}. The first-order
extended Limber approximation provides sub-percent accuracy for $\ell > 10$ and
is sufficient for present surveys. The associated errors are sub-dominant even
for future large surveys.

@@

This paper is organised as follows.  In section~\ref{sec:wl} we provide a
pedagogical introduction to weak gravitational lensing theory, projections and
power spectra for the flat-sky and spherical cases, following the seminal work
\forref{by} \citet{2000PhRvD..62d3007H} \cite[see also][]{2005PhRvD..72b3516C}. In
section~\ref{sec:L2} we derive weak-lensing observables using a second-order
Limber approximation first introduced by \citet{2008PhRvD..78l3506L}. We
compare the shear power spectrum and the commonly-used two-point shear
correlation function for the full solution to a range of different
approximations in section~\ref{sec:results}, providing cosmological constraints
for each case using CFHTLenS data from \citet{CFHTLenS-2pt-notomo}. This paper
draws from several sources of literature which have previously discussed the
full projection, or reviewed the accuracy of the Limber approximation for weak
lensing studies, namely \citet{2008PhRvD..78d3002S,2012PhRvD..86b3001B,
2012MNRAS.422.2854G, 2016arXiv161104954K}\forref{, see also
\cite{2017arXiv170401054L} for a more recent work.} We present a discussion and
comparison of our results to these papers in Appendix~\ref{app:B}.

@@

The lensing potential $\psi$ at a position on the sky $(\theta, \varphi)$ in
the Born approximation is defined as the projected 3D metric potential $\Phi$
along the line of sight of a flat Universe \citep{1998ApJ...498...26K,BS01},
%
% Hu00 (21,22) with minus sign
%
\begin{equation}
  \psi(\theta, \varphi) = \frac 2 {{\rm c}^2} \int\limits_0^\infty \frac{{\rm d}\chi}{\chi}
    %\Phi[\chi, f_K(\chi) \theta, f_K(\chi) \varphi; \chi] \, q(\chi),
    \Phi[\chi, \chi \theta, \chi \varphi; \chi] \, q(\chi),
  \label{eq:psi}
\end{equation}
%
where the lensing efficiency $q$ is given by
%
\begin{equation}
  q(\chi) = \int\limits_\chi^{\chi_{\rm h}} {\rm d} \chi^\prime \, n(\chi^\prime)
    %\frac{f_K(\chi^\prime - \chi)}{f_K(\chi^\prime)},
    \frac{\chi^\prime - \chi}{\chi^\prime},
  %
  \label{eq:lens_efficiency}
\end{equation}
%
corresponding to a population of lensed galaxies with a normalised source
redshift distribution $n_z(z) {\rm d}z = n(\chi) {\rm d} \chi$, with the limit
being the comoving distance to the horizon $\chi_{\rm h}$\footnote{In
(\ref{eq:psi}) we have replaced this limit without loss of generality with
$\infty$, since $q(\chi > \chi_{\rm h}) = 0$.}. Here, ${\rm c}$ is the speed of
light, and the projection is carried out over comoving distances $\chi$.
The last argument of the potential $\Phi$ is not to
be understood as coordinate, but rather as a substitute of cosmic time,
$t(\chi)$, to express the time at which the potential is evaluated. This is
true in the following for all fields and functions thereof that dynamically
change with cosmic epoch.

@@

Following \cite{2000PhRvD..62d3007H} we decompose the lensing
potential (equation~\ref{eq:psi}) into spherical harmonics, in analogy
to the cosmic microwave background (CMB)
temperature, both of which are scalar functions on the sphere. This
decomposition and its inverse are
%
% Hu00 (23)
%
\begin{align}
  \psi(\theta, \varphi) = \sum_{\ell=0}^\infty \sum_{m=-\ell}^\ell \psi_{\ell m} {\rm Y}_{\ell m}(\theta, \varphi);
    \label{eq:psi_harm_exp}
    \\
    %
  \psi_{\ell m} = \int_{\mathbb{S}^2} {\rm d} \Omega \, \psi(\theta, \varphi) {\rm Y}^\ast_{\ell m}(\theta, \varphi).
  \label{eq:psi_harm_exp_inv}
\end{align}
%
Complex conjugation is denoted with the superscript $^\ast$.
To specify tomographic redshift bins $i=1\ldots N_z$, we introduce a family
of lensing efficiency functions $q_i$ defined by a corresponding family of
redshift distributions $n_i$ via equation~(\ref{eq:lens_efficiency}). The resulting
lensing potential is denoted by $\psi_{\ell m, i}$. The tomographic power
spectrum of the lensing potential between two redshift bins $i$ and $j$,
$C_{ij}^\psi$ \citep{pee80} is then defined as
%
\begin{equation}
  \left\langle \psi^{}_{\ell m, i} \, \psi^\ast_{\ell^\prime m^\prime, j} \right\rangle
    = \delta_{\ell \ell^\prime} \delta_{m m^\prime} C^\psi_{ij}(\ell) .
  \label{eq:C_ell_psi}
\end{equation}
%
Note that the argument $\ell$ is a discrete integer variable, and is often written as index, $C_\ell$.
Using the properties of the spherical harmonics (see
App.~\ref{sec:derivations_C} for details) the power spectrum can be written as
%
\begin{align}
  C_{ij}^\psi(\ell) = & \frac 8 {c^4 \pi} 
  \int_0^\infty \frac{{\rm d}\chi}{\chi} q_i(\chi)
  \int_0^\infty \frac{{\rm d}\chi^\prime}{\chi^\prime} q_j(\chi^\prime)
  \int {\rm d} k k^2 \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_\Phi(k; \chi, \chi^\prime)
  \label{eq:C_ell_phi_Pphi} \\
  = & \frac 8 \pi \pref^2
  \int_0^\infty \frac{{\rm d}\chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
  \int_0^\infty \frac{{\rm d}\chi^\prime}{\chi^\prime} \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
  \int \frac{{\rm d} k}{k^2} \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_{\rm m}(k; \chi, \chi^\prime) ;
  \label{eq:C_ell_phi_Pm}
\end{align}
%
where ${\rm j}_\ell$ is the spherical Bessel function of order $\ell$. For
convenience we introduce the normalisation constant $\pref$ as
%
\begin{equation}
  \pref = \frac 3 2 \Omega_{\rm m} \left(\frac{H_0} c\right)^2.
  \label{eq:pref}
\end{equation}
%
The first line expresses $C_{ij}^\psi$ in terms of the 3D potential power spectrum $P_\Phi$, defined as
%
\begin{align}
  \left\langle \hat \Phi(\vec k; \chi) \hat \Phi^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_\Phi(k; \chi, \chi^\prime).
    %\nonumber \\
    %= & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) \pref^2 k^{-4} a^{-1}(\chi) a^{-1}(\chi^\prime)
      %P_{\rm m}(k; \chi, \chi^\prime).
  \label{eq:p_phi}
\end{align}
%
The second line uses the 3D matter density power spectrum $P_{\rm m}$, which is defined analogously as
%
\begin{align}
  \left\langle \hat \delta(\vec k; \chi) \hat \delta^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_{\rm m}(k; \chi, \chi^\prime),
  \label{eq:p_m}
\end{align}
%
and is related to $P_\Phi$ via the absolute square of the Poisson equation
(\ref{eq:poisson}).

@@

The majority of cosmic shear analyses have used the predicted lensing power
spectrum approximated on a flat sky, neglecting the sky curvature. This is a
valid approach for past and current survey areas with an extent less than 10
degrees. To account for the sky curvature of the observed data, the shear
correlation functions from observed galaxy ellipticities \forref{are} now routinely
computed using spherical coordinates, since projecting to a Cartesian plane has
been shown to cause significant biases of the two-point funtion on large scales
\citep{FSHK08}, and lead to spurious B-modes \citep{asgari/etal:2017}. Here
we examine the effect of sky curvature on the theoretical models of the power
spectrum, and the effect on cosmological parameter inference (see
Sect.~\ref{sec:cfhtlens}).

@@

First, we use the identity of Bessel functions
%
\begin{equation}
  {\rm j}_\ell(x) = \sqrt{\frac{\pi}{2x}} {\rm J}_{\ell + 1/2}(x)
  \label{eq:jJ}
\end{equation}
%
in equation~(\ref{eq:C_ell_full}), where ${\rm J}_\nu$ is the Bessel function of the first kind
and order $\nu$. Next, \cite{2008PhRvD..78l3506L} solve integrals of
the form
%
\begin{equation}
  \int_0^\infty {\rm d} \chi f(\chi) {\rm J}_\nu(k \chi)
  = \int_0^\infty {\rm d} x k^{-1} f(x/k) {\rm J}_\nu(x)
  \label{eq:int_dchi}
\end{equation}
%
by performing a Taylor expansion of an arbitrary \forref{differentiable} function
$f$ around $x = k \chi = \nu = \ell + 1/2$, where the Bessel function has its
approximate maximum.

@@

To further develop equation~(\ref{eq:P_geom_mean}), we divide out the growth factor of
the power spectrum,
%
\begin{equation}
 P_{\rm m}(k, \chi) =: D^2_+(\chi) \tilde P_{\rm m}(k, \chi) \forref{\approx D^2_+(\chi) \tilde P_{\rm m}(k)}.
  \label{eq:P_k_chi_sep_s}
\end{equation}
%
The function $\tilde P_{\rm m}$ is in general not time-independent, except
in the case of a linear matter power spectrum in the absence of massive
neutrinos, and in General Relativity. However, the second-order Limber terms
are expected to be important only for small $\ell$, since for large $\ell$ the first-order
Limber equation (\ref{eq:C_ell_limber1}) is dominant. Eq.~(\ref{eq:P_k_chi_sep_s})
becomes a good approximation for small $\ell$, since that means either small
$k$ where the evolution is linear, or small $\chi$, where the lensing
efficiency is small. The accuracy of a very nearby tomographic bin with low
mean redshift should be further examined, but this is not the case for
CFHTLenS.

@@

\forref{We note that in all cases the second-order Limber expansion adds power
to the first-order term. In the flat hybrid case, which over-estimates the full
spherical solution, this results in the second-order expansion being less
precise compared to first-order.}

@@

\renewcommand{\baselinestretch}{1.5}
\begin{table}
\begin{centering}
  %
  \caption{\label{tab:cases}The shear power spectrum approximations studied in this paper. `ID' is the label
    used in the text and figures.
    \forref{The sixth (seventh) column indicates the mode $\ell_x$ such that for $\ell \leq \ell_x$, the approximated
    power spectrum is more accurate than $x$, with $x=0.1$ ($0.01$).}
    } 
    %
    \begin{tabular}{|p{0.22\textwidth}|l|c|c|c|c|c|p{0.21\textwidth}}
  \hline
  Case & ID & equation & $p(\ell)$ & $\nu(\ell)$ & \forref{$\ell_{0.1}$} & \forref{$\ell_{0.01}$} & Comment \\ \hline
  %  black
  $1^{\rm st}$-order standard Limber, flat & L1Fl & (\ref{eq:P_ell_limber1})+(i)
    & $1$ & $\ell$ & $4$ & $60$ & Pre-2014 CFHTLenS and DLS \\ \hline
  % blue
  $1^{\rm st}$-order extended Limber, flat & ExtL1Fl & (\ref{eq:P_ell_limber1})+(ii)
    & $\frac{\ell^4}{(\ell+1/2)^4}$ & $\ell + \frac 1 2$ & $20$ & $200$ & Converges only with ${\cal O}(\ell^{-1})$ \\ \hline
  % pink
  $1^{\rm st}$-order extended Limber, hybrid, flat & ExtL1FlHyb & (\ref{eq:P_ell_limber1})+(iii)
    & $1$ & $\ell + \frac 1 2$ & $4$ & $10$ & Post-2014 CFHTLenS, DES-SV and KiDS \\ \hline
  % cyan
  $1^{\rm st}$-order extended Limber, spherical & ExtL1Sph & (\ref{eq:C_ell_limber1})
    & $\frac{\ellbar^2(\ell, 2)}{(\ell+1/2)^4}$ & $\ell+ \frac 1 2$ & $3$ & $12$ & \\ \hline
  % grey
  $2^{\rm st}$-order extended Limber, flat & ExtL2Fl &  (\ref{eq:P_ell_limber1})+(\ref{eq:P_ell_limber2})+(iii)
    & $\frac{\ell^4}{(\ell+1/2)^4}$ 
    & $\ell+\frac 1 2$ & $19$ & $200$ & Converges only with ${\cal O}(\ell^{-1})$ \\ \hline
  % orange
  $2^{\rm st}$-order extended Limber, hybrid, flat & ExtL2FlHyb &  (\ref{eq:P_ell_limber1})+(\ref{eq:P_ell_limber2})+(iii)
    & $1$ 
    & $\ell+ \frac 1 2$ & $5$ & $16$ & Best flat-sky approximation \\ \hline
  % green
  $2^{\rm st}$-order extended Limber, spherical & ExtL2Sph & (\ref{eq:C_ell_limber1})+(\ref{eq:C_ell_limber2_dr_s})
    & $\frac{\ellbar^2(\ell, 2)}{(\ell+1/2)^4}$ & $\ell+\frac 1 2$ & $2$ & $5$ & Best approximation \\ \hline
  %
  Full spherical & FullSph & (\ref{eq:C_ell_full}) &
      - & - & - & - & Correct projection \\ \hline
  %
  \end{tabular}

@@

The majority of cosmic shear analyses to date have adopted real-space
correlation statistics, since these can be measured directly from an observed
galaxy shape catalogue. The baseline quantity is the two-point correlation
function \citep{1991ApJ...370....1M, 1992ApJ...388..272K, BS01}, given \forref{in the flat-sky approximation} by
%
\begin{align}
  \xi_+(\theta) 
  % &
  = \forref{\left\langle \gamma \gamma^\ast \right\rangle(\theta) =} \frac 1 {2\pi} \int {\rm d} \ell \, \ell \, {\rm J}_0(\ell
   \theta)
  P^\gamma(\ell);
  \quad
  %\nonumber \\
   %
   \xi_-(\theta)
  %&
  = \forref{\left\langle \gamma \gamma \right\rangle(\theta) =} \frac 1 {2\pi} \int
   {\rm d} \ell \, \ell \, {\rm J}_4(\ell \theta)
  P^\gamma(\ell).
  %
   \label{eqn:xiGG}
\end{align}
%
The flat-sky shear power spectrum $P^\gamma$ can be related to the underlying
matter power spectrum through equation~(\ref{eq:P_ell_limber1}) when adopting a
first-order extended Limber approximation, or equation~(\ref{eq:P_ell_limber2})
when adopting a second-order extended Limber approximation.

On a sphere, correlation functions formally cannot be related to the power
spectrum by the Hankel transform in equation~(\ref{eqn:xiGG}), and should be
replaced by the spherical transform
\forref{\citep{1999IJMPD...8...61N,2004MNRAS.350..914C}}
%
\forref{
\begin{equation}
 \xi_+(\theta) 
  =  \frac 1{4\pi} \sum\limits_{\ell=2}^\infty (2 \ell + 1) C^\gamma(\ell) {\rm d}^\ell_{2\,2}(\theta); \quad 
 \xi_0(\theta) 
  =  \frac 1{4\pi} \sum\limits_{\ell=2}^\infty (2 \ell + 1) C^\gamma(\ell) {\rm d}^\ell_{2\,-2}(\theta)
  \label{eq:xi_wigner}
\end{equation}
}
%
\forref{where ${\rm d}^\ell_{m\,n}$ are the reduced Wigner ${\rm D}$-matrices, see App.~\ref{app:C} for details
on their numerical calculation.}

@@

  \caption{%
    \forref{The difference of the two-point shear correlation functions $\xi_+$ (\emph{left}) and $\xi_-$ (\emph{right})
    of the ExtL2Sph projection relative to the full spherical case (FullSph). Two cases of the shear correlation function transformation for
    ExtL2Spha are shown,
    the full spherical case (eq.~\ref{eq:xi_wigner}, green solid lines), and the flat-sky Hankel transform (eq.~\ref{eqn:xiGG}, red dashed curves).}
  }

>>>>

\title[Precision cosmic-shear projection]{Precision calculations of the cosmic shear power spectrum projection}

@@

We compute the spherical-sky weak-lensing power spectrum of the shear and
convergence. We discuss various approximations, such as flat-sky, and first- and
second-order Limber equations for the projection. We find that the impact of
adopting these approximations is negligible when constraining cosmological
parameters from current weak lensing surveys.
%
This is demonstrated using data from the Canada-France-Hawaii Telescope Lensing Survey
(CFHTLenS). We find that the reported tension with Planck Cosmic Microwave
Background (CMB) temperature anisotropy results cannot be alleviated.
%
For future large-scale surveys with unprecedented precision, we show that the
spherical second-order Limber approximation will provide sufficient accuracy.
In this case, the cosmic-shear power spectrum is shown to be in agreement with
the full projection at the sub-percent level for $\ell > 3$, with the
corresponding errors an order of magnitude below cosmic variance for all
$\ell$. When computing the two-point shear correlation function, we show that
the flat-sky fast Hankel transformation results in errors below two percent
compared to the full spherical transformation.
%
In the spirit of reproducible research, our numerical implementation of all
approximations and the full projection are publicly available within the
package \textsc{nicaea} at \texttt{http://www.cosmostat.org/software/nicaea}.

@@

In the era of the upcoming large-scale surveys that will provide measurements
of cosmic shear with unprecedented precision, one needs to revisit the
theoretical predictions of the observables to ensure that the accuracy of the
models meet the accuracy of the observations. In this paper, we examine the
widely used Limber approximation for the projected weak-lensing power spectrum.
We consider spherical coordinates and the flat-sky approximation, and compute
the full projection of the lensing power spectrum. The first-order
extended Limber approximation provides sub-percent accuracy for $\ell > 10$ and
is sufficient for present surveys. The associated errors are sub-dominant even
for future large surveys.

@@

This paper is organised as follows.  In section~\ref{sec:wl} we provide a
pedagogical introduction to weak gravitational lensing theory, projections and
power spectra for the flat-sky and spherical cases, following the seminal work
by \citet{2000PhRvD..62d3007H} \cite[see also][]{2005PhRvD..72b3516C}. In
section~\ref{sec:L2} we derive weak-lensing observables using a second-order
Limber approximation first introduced by \citet{2008PhRvD..78l3506L}. We
compare the shear power spectrum and the commonly-used two-point shear
correlation function for the full solution to a range of different
approximations in section~\ref{sec:results}, providing cosmological constraints
for each case using CFHTLenS data from \citet{CFHTLenS-2pt-notomo}. This paper
draws from several sources of literature which have previously discussed the
full projection, or reviewed the accuracy of the Limber approximation for weak
lensing studies, namely \citet{2008PhRvD..78d3002S,2012PhRvD..86b3001B,
2012MNRAS.422.2854G, 2016arXiv161104954K}, see also
\cite{2017arXiv170401054L} for a more recent work. We present a discussion and
comparison of our results to these papers in Appendix~\ref{app:B}.

@@

The lensing potential $\psi$ at a position on the sky $(\theta, \varphi)$ in
the Born approximation is defined as the projected 3D metric potential $\Phi$
along the line of sight of a flat Universe \citep{1998ApJ...498...26K,BS01},
%
% Hu00 (21,22) with minus sign
%
\begin{equation}
  \psi(\theta, \varphi) = \frac 2 {{\rm c}^2} \int\limits_0^\infty \frac{{\rm d}\chi}{\chi}
    %\Phi[\chi, f_K(\chi) \theta, f_K(\chi) \varphi; \chi] \, q(\chi),
    \Phi(\chi, \chi \theta, \chi \varphi; \chi) \, q(\chi),
  \label{eq:psi}
\end{equation}
%
where the lensing efficiency $q$ is given by
%
\begin{equation}
  q(\chi) = \int\limits_\chi^{\chi_{\rm h}} {\rm d} \chi^\prime \, n(\chi^\prime)
    \frac{\chi^\prime - \chi}{\chi^\prime},
  %
  \label{eq:lens_efficiency}
\end{equation}
%
corresponding to a population of lensed galaxies with a normalised source
redshift distribution $n_z(z) {\rm d}z = n(\chi) {\rm d} \chi$, with the limit
being the comoving distance to the horizon $\chi_{\rm h}$\footnote{In
(\ref{eq:psi}) we have replaced this limit without loss of generality with
$\infty$, since $q(\chi > \chi_{\rm h}) = 0$.}. Here, ${\rm c}$ is the speed of
light, and the projection is carried out over comoving distances $\chi$.
The last argument of the potential $\Phi$ is not to
be understood as coordinate, but rather as a substitute of cosmic time,
$t(\chi)$, to express the time at which the potential is evaluated. This is
true in the following for all fields and functions thereof that dynamically
change with cosmic epoch.

@@

Following \cite{2000PhRvD..62d3007H} we decompose the lensing
potential (equation~\ref{eq:psi}) into spherical harmonics, in analogy
to the cosmic microwave background (CMB)
temperature, both of which are scalar functions on the sphere. This
decomposition and its inverse are
%
% Hu00 (23)
%
\begin{align}
  \psi(\theta, \varphi) = \sum_{\ell=0}^\infty \sum_{m=-\ell}^\ell \psi_{\ell m} {\rm Y}_{\ell m}(\theta, \varphi);
    \label{eq:psi_harm_exp}
    \\
    %
  \psi_{\ell m} = \int_{\mathbb{S}^2} {\rm d} \Omega \, \psi(\theta, \varphi) {\rm Y}^\ast_{\ell m}(\theta, \varphi).
  \label{eq:psi_harm_exp_inv}
\end{align}
%
Complex conjugation is denoted with the superscript $^\ast$.
To specify tomographic redshift bins $i=1\ldots N_z$, we introduce a family
of lensing efficiency functions $q_i$ defined by a corresponding family of
redshift distributions $n_i$ via equation~(\ref{eq:lens_efficiency}). The resulting
lensing potential is denoted by $\psi_{\ell m, i}$. The tomographic power
spectrum of the lensing potential between two redshift bins $i$ and $j$,
$C_{ij}^\psi$ \citep{pee80} is then defined as
%
\begin{equation}
  \left\langle \psi^{}_{\ell m, i} \, \psi^\ast_{\ell^\prime m^\prime, j} \right\rangle
    = \delta_{\ell \ell^\prime} \delta_{m m^\prime} C^\psi_{ij}(\ell) .
  \label{eq:C_ell_psi}
\end{equation}
%
Note that the argument $\ell$ is a discrete integer variable, and is often written as index, $C_\ell$.
Using the properties of the spherical harmonics (see
App.~\ref{sec:derivations_C} for details) the power spectrum can be written as
%
\begin{align}
  C_{ij}^\psi(\ell) = & \frac 8 {c^4 \pi} 
  \int_0^\infty \frac{{\rm d}\chi}{\chi} q_i(\chi)
  \int_0^\infty \frac{{\rm d}\chi^\prime}{\chi^\prime} q_j(\chi^\prime)
  \int {\rm d} k k^2 \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_\Phi(k; \chi, \chi^\prime)
  \label{eq:C_ell_phi_Pphi} \\
  = & \frac 8 \pi \pref^2
  \int_0^\infty \frac{{\rm d}\chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
  \int_0^\infty \frac{{\rm d}\chi^\prime}{\chi^\prime} \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
  \int \frac{{\rm d} k}{k^2} \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_{\rm m}(k; \chi, \chi^\prime) ;
  \label{eq:C_ell_phi_Pm}
\end{align}
%
where ${\rm j}_\ell$ is the spherical Bessel function of order $\ell$. For
convenience we introduce the normalisation constant $\pref$ as
%
\begin{equation}
  \pref = \frac 3 2 \Omega_{\rm m} \left(\frac{H_0} c\right)^2.
  \label{eq:pref}
\end{equation}
%
The first line expresses $C_{ij}^\psi$ in terms of the 3D potential power spectrum $P_\Phi$, defined as
%
\begin{align}
  \left\langle \hat \Phi(\vec k; \chi) \hat \Phi^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_\Phi(k; \chi, \chi^\prime).
  \label{eq:p_phi}
\end{align}
%
The second line uses the 3D matter density power spectrum $P_{\rm m}$, which is defined analogously as
%
\begin{align}
  \left\langle \hat \delta(\vec k; \chi) \hat \delta^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_{\rm m}(k; \chi, \chi^\prime),
  \label{eq:p_m}
\end{align}
%
and is related to $P_\Phi$ via the absolute square of the Poisson equation
(\ref{eq:poisson}).

@@

The majority of cosmic shear analyses have used the predicted lensing power
spectrum approximated on a flat sky, neglecting the sky curvature. This is a
valid approach for past and current survey areas with an extent less than 10
degrees. To account for the sky curvature of the observed data, the shear
correlation functions from observed galaxy ellipticities are now routinely
computed using spherical coordinates, since projecting to a Cartesian plane has
been shown to cause significant biases of the two-point funtion on large scales
\citep{FSHK08}, and lead to spurious B-modes \citep{asgari/etal:2017}. Here
we examine the effect of sky curvature on the theoretical models of the power
spectrum, and the effect on cosmological parameter inference (see
Sect.~\ref{sec:cfhtlens}).

@@

First, we use the identity of Bessel functions
%
\begin{equation}
  {\rm j}_\ell(x) = \sqrt{\frac{\pi}{2x}} {\rm J}_{\ell + 1/2}(x)
  \label{eq:jJ}
\end{equation}
%
in equation~(\ref{eq:C_ell_full}), where ${\rm J}_\nu$ is the Bessel function of the first kind
and order $\nu$. Next, \cite{2008PhRvD..78l3506L} solve integrals of
the form
%
\begin{equation}
  \int_0^\infty {\rm d} \chi f(\chi) {\rm J}_\nu(k \chi)
  = \int_0^\infty {\rm d} x k^{-1} f(x/k) {\rm J}_\nu(x)
  \label{eq:int_dchi}
\end{equation}
%
by performing a Taylor expansion of an arbitrary differentiable function
$f$ around $x = k \chi = \nu = \ell + 1/2$, where the Bessel function has its
approximate maximum.

@@

To further develop equation~(\ref{eq:P_geom_mean}), we divide out the growth factor of
the power spectrum,
%
\begin{equation}
 P_{\rm m}(k, \chi) =: D^2_+(\chi) \tilde P_{\rm m}(k, \chi) \approx D^2_+(\chi) \tilde P_{\rm m}(k).
  \label{eq:P_k_chi_sep_s}
\end{equation}
%
The function $\tilde P_{\rm m}$ is in general not time-independent, except
in the case of a linear matter power spectrum in the absence of massive
neutrinos, and in General Relativity. However, the second-order Limber terms
are expected to be important only for small $\ell$, since for large $\ell$ the first-order
Limber equation (\ref{eq:C_ell_limber1}) is dominant. Eq.~(\ref{eq:P_k_chi_sep_s})
becomes a good approximation for small $\ell$, since that means either small
$k$ where the evolution is linear, or small $\chi$, where the lensing
efficiency is small. The accuracy of a very nearby tomographic bin with low
mean redshift should be further examined, but this is not the case for
CFHTLenS.

@@

We note that in all cases the second-order Limber expansion adds power
to the first-order term. In the flat hybrid case, which over-estimates the full
spherical solution, this results in the second-order expansion being less
precise compared to first-order.

@@

\renewcommand{\baselinestretch}{1.5}
\begin{table}
\begin{centering}
  %
  \caption{\label{tab:cases}The shear power spectrum approximations studied in this paper. `ID' is the label
    used in the text and figures.
    The sixth (seventh) column indicates the mode $\ell_x$ such that for $\ell \leq \ell_x$, the approximated
    power spectrum is more accurate than $x$, with $x=0.1$ ($0.01$).
    } 
    %
    \begin{tabular}{|p{0.22\textwidth}|l|c|c|c|c|c|p{0.21\textwidth}}
  \hline
  Case & ID & equation & $p(\ell)$ & $\nu(\ell)$ & $\ell_{0.1}$ & $\ell_{0.01}$ & Comment \\ \hline
  %  black
  $1^{\rm st}$-order standard Limber, flat & L1Fl & (\ref{eq:P_ell_limber1})+(i)
    & $1$ & $\ell$ & $4$ & $60$ & Pre-2014 CFHTLenS and DLS \\ \hline
  % blue
  $1^{\rm st}$-order extended Limber, flat & ExtL1Fl & (\ref{eq:P_ell_limber1})+(ii)
    & $\frac{\ell^4}{(\ell+1/2)^4}$ & $\ell + \frac 1 2$ & $20$ & $200$ & Converges only with ${\cal O}(\ell^{-1})$ \\ \hline
  % pink
  $1^{\rm st}$-order extended Limber, hybrid, flat & ExtL1FlHyb & (\ref{eq:P_ell_limber1})+(iii)
    & $1$ & $\ell + \frac 1 2$ & $4$ & $10$ & Post-2014 CFHTLenS, DES-SV and KiDS \\ \hline
  % cyan
  $1^{\rm st}$-order extended Limber, spherical & ExtL1Sph & (\ref{eq:C_ell_limber1})
    & $\frac{\ellbar^2(\ell, 2)}{(\ell+1/2)^4}$ & $\ell+ \frac 1 2$ & $3$ & $12$ & \\ \hline
  % grey
  $2^{\rm st}$-order extended Limber, flat & ExtL2Fl &  (\ref{eq:P_ell_limber1})+(\ref{eq:P_ell_limber2})+(iii)
    & $\frac{\ell^4}{(\ell+1/2)^4}$ 
    & $\ell+\frac 1 2$ & $19$ & $200$ & Converges only with ${\cal O}(\ell^{-1})$ \\ \hline
  % orange
  $2^{\rm st}$-order extended Limber, hybrid, flat & ExtL2FlHyb &  (\ref{eq:P_ell_limber1})+(\ref{eq:P_ell_limber2})+(iii)
    & $1$ 
    & $\ell+ \frac 1 2$ & $5$ & $16$ & Best flat-sky approximation \\ \hline
  % green
  $2^{\rm st}$-order extended Limber, spherical & ExtL2Sph & (\ref{eq:C_ell_limber1})+(\ref{eq:C_ell_limber2_dr_s})
    & $\frac{\ellbar^2(\ell, 2)}{(\ell+1/2)^4}$ & $\ell+\frac 1 2$ & $2$ & $5$ & Best approximation \\ \hline
  %
  Full spherical & FullSph & (\ref{eq:C_ell_full}) &
      - & - & - & - & Correct projection \\ \hline
  %
  \end{tabular}

@@

The majority of cosmic shear analyses to date have adopted real-space
correlation statistics, since these can be measured directly from an observed
galaxy shape catalogue. The baseline quantity is the two-point correlation
function \citep{1991ApJ...370....1M, 1992ApJ...388..272K, BS01}, given in the flat-sky approximation by
%
\begin{align}
  \xi_+(\theta) 
  % &
  = \left\langle \gamma \gamma^\ast \right\rangle(\theta) = \frac 1 {2\pi} \int {\rm d} \ell \, \ell \, {\rm J}_0(\ell
   \theta)
  P^\gamma(\ell);
  \quad
  %\nonumber \\
   %
   \xi_-(\theta)
  %&
  = \left\langle \gamma \gamma \right\rangle(\theta) = \frac 1 {2\pi} \int
   {\rm d} \ell \, \ell \, {\rm J}_4(\ell \theta)
  P^\gamma(\ell).
  %
   \label{eqn:xiGG}
\end{align}
%
The flat-sky shear power spectrum $P^\gamma$ can be related to the underlying
matter power spectrum through equation~(\ref{eq:P_ell_limber1}) when adopting a
first-order extended Limber approximation, or equation~(\ref{eq:P_ell_limber2})
when adopting a second-order extended Limber approximation.

On a sphere, correlation functions formally cannot be related to the power
spectrum by the Hankel transform in equation~(\ref{eqn:xiGG}), and should be
replaced by the spherical transform
\citep{1999IJMPD...8...61N,2004MNRAS.350..914C}
%
\begin{equation}
 \xi_+(\theta) 
  =  \frac 1{4\pi} \sum\limits_{\ell=2}^\infty (2 \ell + 1) C^\gamma(\ell) {\rm d}^\ell_{2\,2}(\theta); \quad 
 \xi_-(\theta) 
  =  \frac 1{4\pi} \sum\limits_{\ell=2}^\infty (2 \ell + 1) C^\gamma(\ell) {\rm d}^\ell_{2\,-2}(\theta)
  \label{eq:xi_wigner}
\end{equation}
%
where ${\rm d}^\ell_{m\,n}$ are the reduced Wigner ${\rm D}$-matrices, see App.~\ref{app:C} for details
on their numerical calculation.

@@

  \caption{%
    The difference of the two-point shear correlation functions $\xi_+$ (\emph{left}) and $\xi_-$ (\emph{right})
    of the ExtL2Sph projection relative to the full spherical case (FullSph). Two cases of the shear correlation function transformation for
    ExtL2Spha are shown,
    the full spherical case (eq.~\ref{eq:xi_wigner}, green solid lines), and the flat-sky Hankel transform (eq.~\ref{eqn:xiGG}, red dashed curves).
  }
